---
title: "simpleRegress"
author: "Davy D"
date: "Spring 2024"
output: html_document
---
## Compare all models 
Predict LENGTH based on the variables other than LENGTH. Then order them from worst to best by its value of Adjusted-Rsquared (bigger is better), or by RSE (residual standard error) where smaller is better.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## make a simple regression model

```{r}
bears <- read.csv(file="BEARS.csv",header = TRUE)
#adjust factors to be factors
bears$SEX<- as.factor(bears$SEX)
bears$MONTH<- as.factor(bears$MONTH)
```

## Model LENGTH ~ NECK  this is an example  

```{r}
simp <- lm(LENGTH~NECK,data=bears)
plot(LENGTH~NECK,data=bears)
abline(simp)
summary.lm(simp)
```

This results of this graph show the variable of the NECK has a the effectiveness of measuring finding length of a bear by an adjusted r-squared value of 0.75.

```{r}
simpAGE <- lm(LENGTH~AGE,data=bears)
plot(LENGTH~AGE,data=bears)
abline(simpAGE)
summary.lm(simpAGE)
```

This results of this graph show the variable of the AGE  has a the effectiveness of measuring finding length of a bear by an adjusted r-squared value of 0.50.




```{r}
simpHEADLEN <- lm(LENGTH~HEADLEN,data=bears)
plot(LENGTH~HEADLEN,data=bears)
abline(simpHEADLEN)
summary.lm(simpHEADLEN)
```

This results of this graph show the variable of the HEADLENGTH has a the effectiveness of measuring finding length of a bear by an adjusted r-squared value of 0.84.

```{r}
simpHEADWTH <- lm(LENGTH~HEADWTH,data=bears)
plot(LENGTH~HEADWTH,data=bears)
abline(simpHEADWTH)
summary.lm(simpHEADWTH)
```

This results of this graph show the variable of the HEADWIDTH  has a the effectiveness of measuring finding length of a bear by an adjusted r-squared value of 0.53.

```{r}
simpCHEST <- lm(LENGTH~CHEST,data=bears)
plot(LENGTH~CHEST,data=bears)
abline(simpCHEST)
summary.lm(simpCHEST)
```

This results of this graph show the variable of the CHEST has a the effectiveness of measuring finding length of a bear by an adjusted r-squared value of 0.78.

```{r}
simpWEIGHT <- lm(LENGTH~WEIGHT,data=bears)
plot(LENGTH~WEIGHT,data=bears)
abline(simpWEIGHT)
summary.lm(simpWEIGHT)
```

This results of this graph show the variable of the CHEST has a the effectiveness of measuring finding length of a bear by an adjusted r-squared value of 0.74.

The following graphs can be ranked from as follows; The worst at being able to determine the length of a bear are
AGE with the R-squared value of 0.50
HEADWIDTH with the R-squared value of 0.53
WEIGHT with the R-squared value of 0.74
CHEST with the R-squared value of 0.78
HEAD LENGTH with the R-squared value of 0.84 

Out of all the compared models the best simple model so far to find the length of a bear is LENGTH~ HEADLENGTH. With the highest Adjusted R-squared value of 0.84.

